shrimp_goby_boatnoise

### 2025-03-04 ms_boatnoise_shrimp_goby.Rmd
Updated the abstract

**SUMMARY**  
Human activities are reshaping marine soundscapes, with boat noise a near-constant source of underwater noise and a primary contributor to rising oceanic noise pollution. Despite increasing concern, empirical data for most marine life, particularly in natural settings, remains sparse. Fish and crustaceans are underrepresented in noise pollution research, despite their auditory ranges make them especially susceptible to low-frequency typical of boat engines. Accordingly, this study employs a field-based experimental approach to investigate the impacts of boat noise on an iconic behavioural mutualism between gobies (Amblyeleotris steinitzi) and shrimp (Alpheus spp). Our findings demonstrate that both species exhibit increased burrow use in response to boat noise, suggesting a heightened perception of risk. Despite the perceived risk triggered by boat noise, the behaviour of the two species remained tightly correlated and communication—measured as antennae contact—remained unchanged. Notably, shrimp exhibited stronger responses to 2-stroke engines, whereas gobies responded more strongly to 4-stroke engines, indicating species-specific sensitivities to different noise spectra. These results highlight the complex and taxon-specific effects of noise pollution on marine species and emphasize the need for targeted research to inform conservation strategies aimed at mitigating anthropogenic acoustic disturbances in marine ecosystems .

>>>>

**SUMMARY**  
Human activities are reshaping marine soundscapes globally, with persistent low-frequency noise from boat traffic recognised as a primary driver of marine noise pollution. Despite growing concern, empirical data for most marine life, in natural settings, remains sparse, particularly for fish and crustaceans whose auditory ranges overlap with the frequencies of boat noise. Accordingly, we conducted a field experiment to evaluate how boat noise from different engine types (2-stroke and 4-stroke) affects the mutualistic partnership between Steinitz’s goby (Amblyeleotris steinitzi) and snapping shrimp (Alpheus spp.). Across 123 mutualistic partnerships, we recorded behavioural responses before, during, and after noise exposure. Both species increased burrow use during noise exposure, indicating heightened risk perception, yet their tactile communication—critical for the functioning of the mutualism—remained unaltered. Notably, gobies responded more strongly to 4-stroke boat noise, reducing time outside their burrows by 22%, while shrimp were more sensitive to 2-stroke noise, reducing burrow use by 34%, reflecting species-specific sensitivities to different noise spectra. Behavioural changes were transient, rapidly reverting to baseline levels after the noise exposure stopped. These findings highlight divergent vulnerabilities tied to different engine acoustics and emphasise the need for targeted research to inform strategies for mitigating marine noise pollution.
